---
title: "Data Analysis and Statistical Inference"
author: "LKB"
output: 
  html_document: 
    highlight: pygments
    theme: cerulean
---

```{r, echo=FALSE}
setwd("d:/tmp/Dropbox/Edu/Coursea/DataAnalysisAndStatisticalInference/")
rm(list=ls(all=TRUE))

library(knitr)
opts_chunk$set(echo = TRUE, cache = T, cache.path = "cache/", fig.path = "figure/", warning = FALSE)
#http://yihui.name/knitr/options/
```

#Binormal distribution

Suppose there are twelve multiple choice questions in an English class quiz. Each question has five possible answers, and only one of them is correct. Find the probability of having four or less correct answers if a student attempts to answer every question at random.

```{r}
#exactly 4 correct
dbinom(4, size=12, prob=0.2) 

#4 or less
pbinom(4, size=12, prob=0.2) 
```

So following cat question:

## Cats on YouTube
If you randomly select a video on YouTube, the probability that it involves a cat is 0.11. Over the course of a week, you watch 100 videos on YouTube using an app that randomly selects videos (the random video picker). How many cat videos would you need to see to suspect that the random video picker is biased towards cat videos (i.e. yields an unusually high number of cat videos compared to what is expected)?


The correct answer will be probablity of sucess << .11

```{r}
#4 or less
pbinom(4, size=100, prob=0.11) 

#7 or less
pbinom(7, size=100, prob=0.11) 

#18 or more
1 - pbinom(18, size=100, prob=0.11) 


#31 or more
1 - pbinom(31, size=100, prob=0.11) 
```

So its 31+, which is also logical.


##Blood testing. 

In a blood testing procedure, blood samples from 4 people are combined into one mixture. The mixture will only test negative if all the individual samples are negative. If the probability that an individual sample tests positive is 0.12, what is the probability that the mixture will test positive? Assume blood samples are independent of each other.

```{r}
#we will only fail if exactly 4 incorrect
1-dbinom(0, size=4, prob=0.12) 

#eqivalent to
1 - (1-.12)^4
```


#Stats

A 2013 poll found that 24% of 3,226 randomly sampled Chinese adults see crime as a serious issue. Suppose an issue will get special attention from the Chinese government if more than 1-in-5 Chinese adults express concern on an issue.

Suppose you want to conduct this hypothesis test using a simulation method. Which of the following is the most appropriate set up for simulating the null distribution, and calculating the p-value for this hypothesis test?

* *Place 100 chips in a bag, 24 red and 76 green. Sample, with replacement, 3226 chips, and record the proportion of reds. Repeat many times, and calculate the proportion of simulations where the proportion of reds is at least 20%.*

Suppose you want to conduct the earlier hypothesis test using a Central Limit Theorem based ("theoretical") method. Suppose also that you want to construct a confidence interval for the proportion of Chinese adults who see crime as a serious issue using this same sample. What standard error(s) should you use for the hypothesis test (HT) and the confidence interval (CI)?