---
title: "Data Analysis and Statistical Inference Lab 6 -  Linear regression"
author: "LKB"
output: 
  html_document: 
    highlight: tango
    theme: readable
---

```{r, echo=FALSE}
setwd("d:/tmp/Dropbox/Edu/Coursea/DataAnalysisAndStatisticalInference/")
rm(list=ls(all=TRUE))

library(knitr)
opts_chunk$set(echo = TRUE, cache = FALSE, cache.path = "cache/", fig.path = "figure/", warning = FALSE)
#http://yihui.name/knitr/options/
```

#Batter up


```{r}
load(url("http://www.openintro.org/stat/data/mlb11.RData"))

colnames(mlb11)
str(mlb11)
```

In addition to runs scored, there are seven traditionally used variables in the data set: at_bats, hits, homeruns, bat_avg (batting average), strikeouts, stolen_bases, and wins. Also see [Rules of baseball](https://en.wikipedia.org/wiki/Baseball_rules) and a [description of these statistics](, visit  and http://en.wikipedia.org/wiki/Baseball_statistics).


```{r}
require(stats)

plot(mlb11$runs,mlb11$bat_avg)

text <- sprintf("Correlation is %.2f",cor(mlb11$runs, mlb11$at_bats))
text(600,.27, text, col = "blue", font = 4)
abline(lsfit(mlb11$runs,mlb11$bat_avg), col = "blue", lwd = 3, lty = 2)



```




#Sum of squared residuals


```{r}
plot_ss(x = mlb11$at_bats, y = mlb11$runs)
plot_ss(x = mlb11$at_bats, y = mlb11$runs, showSquares = TRUE)
```

The relationship is positive, linear, and moderately strong. One of the potential outliers is a team with approximately 5520 at bats.

#The linear model

```{r}
m1 <- lm(runs ~ at_bats, data = mlb11)
summary(m1)
```

```{r}
m1 <- lm(runs ~ homeruns, data = mlb11)
summary(m1)

plot(mlb11$runs,mlb11$homeruns)
abline(lsfit(mlb11$runs,mlb11$homeruns), col = "blue", lwd = 3, lty = 2)

```

For each additional home run, the model predicts 1.83 more runs, on average.

#Prediction and prediction errors


```{r}
m1 <- lm(runs ~ at_bats, data = mlb11)
plot(mlb11$runs ~ mlb11$at_bats)
abline(m1, col = "blue", lwd = 3, lty = 2)
mlb11$runs[mlb11$at_bats == 5579]
```

#Model diagnostics

To assess whether the linear model is reliable, we need to check for 

1. linearity
2. nearly normal residuals
3. constant variability.


```{r}
plot(m1$residuals ~ mlb11$at_bats) #linearity
abline(h = 0, col = "red", lwd = 2, lty = 3) 

hist(m1$residuals)  #nearly normal residuals


qqnorm(m1$residuals) #constant variability
qqline(m1$residuals, col = "red", lwd = 2, lty = 3) 

```


* The residuals appear to be randomly distributed around 0.
* The residuals are fairly symmetric, with only a slightly longer tail on the right, hence it would be appropriate to deem the the normal distribution of residuals condition met.

```{r}
m1 <- lm(runs ~ at_bats, data = mlb11)
summary(m1)
m1 <- lm(runs ~ hits, data = mlb11) 
summary(m1)
m1 <- lm(runs ~ wins, data = mlb11) 
summary(m1)
m1 <- lm(runs ~ bat_avg, data = mlb11)
summary(m1)
```

```{r}
m1 <- lm(runs ~ new_obs, data = mlb11) 
summary(m1)
m1 <- lm(runs ~ new_slug, data = mlb11) 
summary(m1)
m1 <- lm(runs ~ new_onbase, data = mlb11)
summary(m1)
```