---
title: "Data Analysis and Statistical Inference Lab2"
author: "LKB"
output: html_document
---

```{r, echo=FALSE}
setwd("d:/tmp/Dropbox/Edu/Coursea/DataAnalysisAndStatisticalInference/")

```


#Unit 2 Lab 2 -  Probability

lets investigate hot hands phenomena in basketball

```{r}
load(url("http://www.openintro.org/stat/data/kobe.RData"))
head(kobe)
kobe$basket[1:9]
```

##Q3
Which of the following is false about the distribution of Kobe's streak lengths from the 2009 NBA finals. 
```{r}
kobe_streak <- calc_streak(kobe$basket)
barplot(table(kobe_streak))
boxplot(kobe_streak)
summary(kobe_streak)
```

##Simulations in R
While we don't have any data from a shooter we know to have independent shots, that is very easy to simulate in R. In a simulation, you set the ground rules of a random process and then the computer uses random numbers to generate an outcome that adheres to those rules. As a simple example, you can simulate flipping a fair coin with the following.


```{r}
outcomes <- c("heads", "tails")
sim_fair_coin <-sample(outcomes, size = 100, replace = TRUE)
table(sim_fair_coin)
```

Say we're trying to simulate an unfair coin that we know only lands heads 20% of the time
```{r}
sim_unfair_coin <- sample(outcomes, size = 100, replace = TRUE, prob = c(0.2, 0.8))
table(sim_unfair_coin)
```

```{r}
#?sample #will open another browser window
```

##Simulating the Independent Shooter
Simulating a basketball player who has independent shots uses the same mechanism that we use to simulate a coin flip.

```{r}
library(scales)
outcomes <- c("H", "M")
sim_Kobe <- sample(outcomes, size = length(kobe$basket), replace = TRUE,prob = c(0.45,0.55))

kobe_streak_sim <- calc_streak(sim_Kobe)
table(kobe_streak_sim)
summary(kobe_streak_sim)

#kobe resuls
kobe_streak <- calc_streak(kobe$basket)
table(kobe_streak)
#boxplot(kobe_streak,add=T,col='red')
summary(kobe_streak)

#visuals
barplot(table(kobe_streak_sim),col='skyblue')
barplot(table(kobe_streak),add=T,col=scales::alpha('red',.5))

```

for ggplot try <http://stackoverflow.com/questions/3541713/how-to-plot-two-histograms-together-in-r>

