---
title: "Data Analysis and Statistical Inference"
#author: "LKB"
output: 
  html_document: 
    highlight: tango
    theme: readable
---

```{r, echo=FALSE}
setwd("d:/tmp/Dropbox/Edu/Coursea/DataAnalysisAndStatisticalInference/")
rm(list=ls(all=TRUE))

library(knitr)
opts_chunk$set(echo = TRUE, cache = FALSE, cache.path = "cache/", fig.path = "figure/", warning = FALSE)
#http://yihui.name/knitr/options/
```


#Introduction

Global warming is an important issue which might affect future generation. Scientific community agrees that the greenhouse effect is responsible for global warming with discussion focused on to what extend human activity is changing this process [2].

Unfortunately over the last decade, this topic has been highly politicized, with political lobbing overtaking rational discussion. Given that, it is possible that interest in politics will translate to polarised anty-global warming views. Such influence creates a dangerous situation and preventing discussion. Therefore I want to answer the following research question:


*"Does interest in politics influence ones perception regarding the global warming."*

#Data collection

To answer this question I will be using **American National Elections Study, 2012 Time Series Study Coursera Extract** [1].  I will be looking at following variables:

* INTEREST_ATTENTION answering question *How often do you pay attention to what's going on in government and politics?*
* ENVIR_GWARM answering question *You may have heard about the idea that the world's temperature may have been going up slowly over the past 100 years. What is your personal opinion on this?* 

Both variables are categorical. This is the observational study, as correspondents have been surveyed about their current opinion. A random sampling, based on address have been conducted on sample of USA citizens 18 years or older.  Data have been collected both in face to face interviews and on-line data collection.

Given that I believe that my findings can be generalised to the population. As study was voluntary, there is possibility of voluntary response bias, showing results more strongly opinionated than general population. To reduce this bias I will limit my analysis to people who answered both parts of the study (pre and post election). 

This is observation data, so casual links between the variables of interest can't be established.

As discussed in next section, we have a total of 5384 cases - responses from the correspondents (face to face or online).

#Exploratory data analysis

I will only consider:

* people who responded to both Pre-Election And Post-Election Interviews
* people who gave answers to both questions (removed 126 not complete cases)

```{r}
load(url("http://bit.ly/dasi_anes_data"))

columns2Keep = c('caseid','interest_attention','envir_gwarm')

anes = anes[anes$wave_completions == 'Yes (Pre-Election And Post-Election Interviews Completed)',columns2Keep]
anes <- anes[complete.cases(anes),] #remove participants with any of data missing
str(anes)
summary(anes) 

```

Overall we have `r length(anes$interest_attention)` cases - responses from the correspondents (face to face or online).

Lets visualise this data.

```{r}
globalWarmingBelievers = table(anes$interest_attention,anes$envir_gwarm)

mosaicplot(globalWarmingBelievers, main ="Do you believe in global warming?", xlab = "Do you pay attention to what's going on in government and politics?",ylab = "Is global warming happening?", color = 3:4)

```

Intial exploration of the data suggest that apart from extreme views (Never/Always) consensus is that global warming has been happening.

#Inference

Following my reserach question *"Does interest in politics influence ones perception regarding the global warming"*, I will state following hipothesis

$H_0:$ Interest in politics does not affect ones views about global warming and proportion of people beliving in global warming is constant across all groups.
$H_A:$ Interest in politics influence ones views about global warming.

To prove my theory I will run $\chi^2$ test of independence, thesing if selected categorical variables are dependent of each other.

##Check conditions

In order to use $\chi^2$ test of independence sample has to be:

* independent - test description indicate that samples (cases) have been random, consist <10% of population and each case is counted only once
* each scenario (interest in politics) nave at least 5 expected cases

##Results

```{r}
source("http://bit.ly/dasi_inference")

inference(y = anes$envir_gwarm, x = anes$interest_attention, est = "proportion", type = "ht",  alternative = "greater", method = "theoretical",eda_plot = FALSE, inf_plot = FALSE )

```

Results indicate that Response and explanatory variable are dependent. Following analysis of the data in the previos section I can assume that this is mostly to responses from extreme views (Never/Always) and otherwise citizens behaviour is not affected.

##Do extreme views (Never/Always) skew results?

```{r}
source("http://bit.ly/dasi_inference")

allowedResponses = levels(anes$interest_attention)
allowedResponses = allowedResponses[2:4] #remove extreme responses

nonExtremeCorrespondents = anes[anes$interest_attention %in% allowedResponses,]
nonExtremeCorrespondents$interest_attention = factor(nonExtremeCorrespondents$interest_attention) #remove dropped levels

inference(y = nonExtremeCorrespondents$envir_gwarm, x = nonExtremeCorrespondents$interest_attention, est = "proportion", type = "ht",  alternative = "greater", method = "theoretical",eda_plot = FALSE, inf_plot = FALSE )

USAConsensus= table(nonExtremeCorrespondents$envir_gwarm)/length(nonExtremeCorrespondents$envir_gwarm)
```

Without extreme cases there is no evidence to reject $H_0$ hypothesis.


#Conclusion

In this study we explored the reserach question *"Does interest in politics influence ones perception regarding the global warming"*. 
Results indicate, that apart from extreme cases American society in general (`r USAConsensus[1]*100`%) believes in global warming. 
This proportion is different for two groups with very strong belives (Always/Never). Interestingly proportion of citizens not beliving in global warming is highest in the latest group.
Those results indicate that there is no evidence of interest in politics affecting understanding of scientific research. What would be interesting is to understand why such large number of citizens not following politics is does not believe in global warming.


#References

[1] American National Election Studies (ANES; www.electionstudies.org). The ANES 2012 Time Series Study [dataset]. Stanford University and the University of Michigan [producers].

Further information about the dataset can be found at <https://d396qusza40orc.cloudfront.net/statistics%2Fproject%2Fanes1.html#envir_gwarm>. Full description of study can be accessed at <http://electionstudies.org/studypages/anes_timeseries_2012/anes_timeseries_2012_userguidecodebook.pdf>

[2] Silver, Nate (2012-09-27). The Signal and the Noise: The Art and Science of Prediction (p. 376). Penguin Books Ltd. Kindle Edition have been used.


#Appendix

Attached is the sample of my final dataset

```{r}
head(anes,40)
```

